---
title: "HW3_yx2857"
author: "Yifan Xu"
date: "2024-10-13"
output: github_document
---

```{r setup, include=FALSE}
library(tidyverse)
library(ggridges)
library(patchwork)

knitr::opts_chunk$set(
	echo = TRUE,
	warning = FALSE,
	fig.width = 8, 
  fig.height = 6,
  out.width = "90%"
)

theme_set(theme_minimal() + theme(legend.position = "bottom"))

options(
  ggplot2.continuous.colour = "viridis",
  ggplot2.continuous.fill = "viridis"
)

scale_colour_discrete = scale_colour_viridis_d
scale_fill_discrete = scale_fill_viridis_d


```


**Problem 1**  
```{r}
library(p8105.datasets)
data("ny_noaa")

ny_noaa %>% summary()

ny_noaa <- ny_noaa %>% 
  separate(date, into = c("year", "month", "day"), convert = TRUE) %>% 
  mutate(
    tmax = as.numeric(tmax),
    tmin = as.numeric(tmin))

ny_noaa %>% 
  count(snow) %>%
  arrange(desc(n))

```

 
The `NY_NOAA` dataset contains `r nrow(ny_noaa)` observations and `r ncol(ny_noaa)` variables. The variables are: `r paste(names(ny_noaa), collapse = ", ")`. The dataset contains information about the weather in New York, and the variables include weather station id, date of observation,  precipitation (tenths of mm), snowfall (mm), snow depth (mm), and min and max temperature (tenths of degrees C).  
Missing data is found cross observations for temperature, precipitation, and snowfall. The missing rate of min and max temperature are the highest, which are `r paste0(round(100*sum(is.na(ny_noaa$tmin))/nrow(ny_noaa),0), "%")` and `r paste0(round(100*sum(is.na(ny_noaa$tmax))/nrow(ny_noaa),0), "%")`, respectively. The missing rate of precipitation is `r paste0(round(100*sum(is.na(ny_noaa$prcp))/nrow(ny_noaa),0), "%")`, the missing rate of snowfall is `r paste0(round(100*sum(is.na(ny_noaa$snow))/nrow(ny_noaa),0), "%")`, and the missing rate of snow depth is `r paste0(round(100*sum(is.na(ny_noaa$snwd))/nrow(ny_noaa),0), "%")`.  

0 is the most commonly observed value for snowfall. This is because most days of the year, it does not snow at all in NY. The second most commonly observed value is `NA`, indicating missingness. Other common values are 25, 13, 51, and 76, suggesting that snowfall is originally recorded in units of inch and converted to mm.  

  

```{r}
ny_noaa  %>% 
  group_by(id, year, month) %>% 
  filter(month %in% c(1, 7)) %>% 
  summarize(mean_tmax = mean(tmax, na.rm = TRUE, color = id)) %>% 
  ggplot(aes(x = year, y = mean_tmax, group = id)) + geom_point() + geom_path() +
  facet_grid(~month) +
  labs(title = "Mean monthly temperature for each station across years for January and July")

```

The mean temperature in January is much lower than the mean temperature in July for all stations and across all years. We can observe that temperature measurements at all stations are highly uniform in trend, from peak to valley, across the years, and this pattern appears both in January and July.  
We can see one uncharacteristically cold station in July of 1987 or 1988, as well as a few other less drastic outliers.  


_Make a two-panel plot showing_   
(i) tmax vs tmin for the full dataset (note that a scatterplot may not be the best option)  
(ii) make a plot showing the distribution of snowfall values greater than 0 and less than 100 separately by year.  

```{r}
hex = 
  ny_noaa %>% 
  ggplot(aes(x = tmin, y = tmax)) + 
  geom_hex()+
  theme(legend.text = element_text(size = 5) )

ridge = 
  ny_noaa %>% 
  filter(snow < 100, snow > 0) %>%
  ggplot(aes(x = snow, y = as.factor(year))) + 
  geom_density_ridges()

hex + ridge
```

```{r}
rm(ny_noaa )

```

**Problem 2**  


```{r}


accel <- read_csv("data/nhanes_accel.csv")
  
covar <- read_csv("data/nhanes_covar.csv", skip = 4) %>% 
        filter(age >= 21)  %>% 
        na.omit %>%
        mutate(sex = factor(sex, levels = c(1, 2), labels = c("Male", "Female"))) %>%
        mutate(education = factor(education, levels = c(1, 2, 3), labels = c("Less than high school", "High school equivalent", "More than high school")))

          
accel_tidy <- left_join(covar, accel) %>% janitor::clean_names()

```


```{r}
# Produce a reader-friendly table for the number of men and women in each education category
accel_tidy %>% 
  count(sex, education) 
```

```{r}
# create a visualization of the age distributions by sex in each education category
accel_tidy %>% ggplot(aes(x = sex, y = age, fill = sex)) +
  geom_boxplot() +
  facet_wrap(~ education) +
  labs(title = "Age Distribution by Sex in Each Education Category",
       x = "Sex",
       y = "Age") 
  
```

From the table and the plot, we can see that male and female participants are evenly distributed across education levels. The age distribution of participants is similar in education category "less than high school" and "More than high school", and median age of female with "High school equivalent" is higher than the male. The median age of participants in the "More than high school" category being slightly lower than the other two categories.  



```{r}
# Traditional analyses of accelerometer data focus on the total activity over the day. Using your tidied dataset, aggregate across minutes to create a total activity variable for each participant. Plot these total activities (y-axis) against age (x-axis); your plot should compare men to women and have separate panels for each education level. Include a trend line or a smooth to illustrate differences. Comment on your plot.

accel_tidy %>% 
   mutate(total_activity = rowSums(select(., starts_with("min")))) %>% 
   ggplot(aes(x = age, y = total_activity, color = sex)) +
   geom_point(alpha = 0.5) + 
   geom_smooth() +
  facet_wrap(~ education) +  # Facet the plot by education level
  labs(
    title = "Total Activity vs. Age by Sex and Education Level",
    x = "Age",
    y = "Total Activity"
  )  

```
According to the plot, we can see that the total activity of participants decreases with age. The total activity of female participants seems to have a steeper decline than male participants in the "Less than high school" category of education. The total activity of female participants with "High school equivalent" and "More than high school" education level is higher than male participants. The overall total activity of participants with "Less than high school" education level at younger age is higher than the other two education levels at younger age.   

```{r}
# Accelerometer data allows the inspection activity over the course of the day. Make a three-panel plot that shows the 24-hour activity time courses for each education level and use color to indicate sex. Describe in words any patterns or conclusions you can make based on this graph; including smooth trends may help identify differences. 



accel_tidy %>%
  pivot_longer(cols = starts_with("min"), names_to = "minute", values_to = "activity") %>%
  # Convert the minute column to numeric, then convert minutes to hours
  mutate(hour = as.numeric(str_remove(minute, "min")) / 60) %>%
  ggplot(aes(x = hour, y = activity, color = sex)) +
  geom_line(alpha = 0.5) +  
  geom_smooth(method = "loess", se = FALSE) +  
  facet_wrap(~ education) + 
  labs(
    title = "24-hour Activity Time Courses by Education Level",
    x = "Hour of the Day",
    y = "Activity Level",
    color = "Sex"
  ) +
  scale_x_continuous(breaks = seq(0, 24, by = 2)) + 
  theme_minimal()

```
From this plot, we can learn the male participants with "More than high school" are having more activity at night time than female of this education. And participants with "more than high school", both male and female are having more activity at early morning time than the other two categories.  

```{r}
rm(accel, accel_tidy, covar)

```

**Problem 3**

```{r, show_col_types = FALSE }
# Citi Bike is a bike sharing system in New York City; riders can rent pedal-powered or electric bikes from starting stations and return them to another station at their destination. Introduced in 2013, the system is immensely popular and expanded rapidly. Riders who use this system often may become “members” with lower rental rates.
# 
# This zip file contains data on rides taken on the NYC Citi Bike system. Files contain 1% of all rides with a total duration less than 4 hours in each of four months. Import, clean, and tidy these data, and describe the resulting dataset.

citi_bike <- read_csv("data/citibike/Jan 2020 Citi.csv") %>% 
  mutate( year = "2020", month = "January") %>%
  bind_rows(read_csv("data/citibike/July 2020 Citi.csv")%>% 
  mutate( year = "2020", month = "July") ) %>% 
  bind_rows(read_csv("data/citibike/Jan 2024 Citi.csv")%>% 
  mutate( year = "2024", month = "January") ) %>% 
  bind_rows(read_csv("data/citibike/July 2024 Citi.csv")%>% 
  mutate( year = "2024", month = "July") ) %>% 
  janitor::clean_names()

# Produce a reader-friendly table showing the total number of rides in each combination of year and month separating casual riders and Citi Bike members. Comment on these results.

citi_bike %>% 
  count(year, month, member_casual) 
```
According to the table, the number of rides in 2024 is higher than 2020. The number of rides in January is lower than July in the same year. The number of rides for members is higher than casual riders, but there are more casual riders in 2024 in terms of proportion.  

```{r}

# Make a table showing the 5 most popular starting stations for July 2024; include the number of rides originating from these stations.

citi_bike %>% 
  filter(month == "July", year == "2024") %>% 
  count(start_station_name) %>% 
  arrange(desc(n)) %>% 
  head(5)
```

```{r}
# Make a plot to investigate the effects of day of the week, month, and year on median ride duration. This plot can include one or more panels, but should facilitate comparison across all variables of interest. Comment on your observations from this plot.
 
citi_bike %>%
  group_by(year, month, weekdays) %>%
  summarize(median_duration = median(duration, na.rm = TRUE), .groups = "drop") %>% 
  ggplot(aes(x = factor(weekdays, levels = c("Monday", "Tuesday", "Wednesday", "Thursday", "Friday", "Saturday", "Sunday")), 
             y = median_duration, 
             color = interaction(year, month),  
             linetype = interaction(year, month),   
             group = interaction(year, month))) +   
  geom_line(size = 1) +  
  labs(title = "Median Ride Duration by Weekdays, Year, and Month",
       x = "Weekdays",
       y = "Median Ride Duration",
       color = "Year-Month",    
       linetype = "Year-Month") +
  theme(legend.position = "right")



```
According to the plot, we can see that the median ride duration is higher in July than Jan. in the same year. The median ride duration is higher in July than January. The median ride duration is generally higher on weekends than weekdays, except for Jan. 2024.  


```{r}

# There were relatively few electric Citi Bikes in 2020, but many more are available now. For data in 2024, make a figure that shows the impact of month, membership status, and bike type on the distribution of ride duration. Comment on your results

citi_bike %>% 
  filter(year == "2024") %>% 
  ggplot(aes(x = duration, fill = member_casual)) + 
  geom_density(alpha = 0.5) + 
  facet_grid(~ month + rideable_type) + 
  labs(title = "Distribution of Ride Duration by Month, Membership Status, and Bike Type",
       x = "Duration (minutes)",
       y = "Density",
       fill = "Membership Status") +
  theme_minimal()

```
  
From the plot, we can read that the distribution of ride duration for casual riders is flatter than members, and the distribution of ride duration for electric bikes is less spread than classic bikes. The shorter ride duration for electronic bike riders is higher in January than July. The distribution of classic ride duration is pretty consistent in Jan. and July.  